Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses

### 2014-06-03 doc/LinksToDAG.tex
minor edits

\documentclass[11pt]{article}
\usepackage{acl2014}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{xspace}

@@

As recently discussed by \newcite{gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
\begin{itemize}
\item {\sc Single-Head}: each word has $\leq 1$ parent
\item {\sc Acyclic}: there are no directed cycles
\item {\sc Connected}: each pair of words has a undirected path between them
\item {\sc Reachable}: each word can be reached from 0 by a directed path
  \noteJE{note that this implies connected?  or does connected not allow 0 words on the path, i.e., does it restrict 0 to one child, and if not, is there a principle for that?  If no one, mention in footnote 1.}
\item {\sc Planar}: edges may not ``cross'' \\ (if there are edges between $u,v$ and between $i,j$, where $u < i < v$, then $u \leq j \leq v$)
\end{itemize}
It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.\footnote{{\sc Projective} basically means {\sc Planar} + {\sc Reachable}.}  However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.

@@

It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.  However, this is not quite true.  Link Grammar \cite{SleatorTemperly91} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.  It produces undirected connected planar graphs.  Annotation conventions are implicit in the detailed lexicon for the Link Grammar Parser,\noteJE{check caps}%
\footnote{\url{http://www.abisource.com/projects/link-grammar/dict/introduction.html}.  The 122 link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html}.} 
which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.  As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.  For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.

@@

>>>>

\documentclass[11pt]{article}
\usepackage{acl2014}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{enumitem}

@@

As recently discussed by \newcite{gomezrodriguez-nivre-2013}, one might impose various requirements on the parse graph:
\begin{itemize}[noitemsep]
\item {\sc Single-Head}: each word has $\leq 1$ parent
\item {\sc Acyclic}: there are no directed cycles
\item {\sc Connected}: each pair of words has a undirected path between them
\item {\sc Reachable}: each word can be reached from 0 by a directed path
  \noteJE{note that this implies connected?  or does connected not allow 0 words on the path, i.e., does it restrict 0 to one child, and if not, is there a principle for that?  If no one, mention in footnote 1.}
\item {\sc Planar}: edges may not ``cross'' \\ (if there are edges between $i,j$ and between $u,v$, where $i < u < j$, then $i \leq v \leq j$)
\end{itemize}
It is common to impose all of these requirements at once, leading to a {\em projective dependency parser} that produces projective trees rooted at 0.\footnote{{\sc Projective} basically means {\sc Planar} + {\sc Reachable}.}  However, parsing algorithms can be devised that relax any or all of the requirements \cite{gomezrodriguez-nivre-2013}.

@@

It seems at first that no one has worked out annotation conventions for {\em projective} multi-headed dependency parsing.  However, this is not quite true.  Link Grammar \cite{SleatorTemperly91} is a grammar-based formalism for projective dependency parsing with {\em undirected} links.  It produces undirected connected planar graphs.  Annotation conventions are implicit in the detailed lexicon for the Link Grammar Parser,\noteJE{check caps}%
\footnote{\url{http://www.abisource.com/projects/link-grammar/dict/introduction.html}.  The 122 link types in the English lexicon are documented at \url{http://www.abisource.com/projects/link-grammar/dict/summarize-links.html}.} 
which specifies for every word a constraint on the {\em sequence} of labeled leftward and rightward edges attached to it.  As remarked by \newcite{eisner-2000-iwptbook}, this is analogous to dependency grammar's use of head automata to constrain a word's sequence of left and right children.  For example, in {\em ``The boy that Jill skipped with fell down,''} the word {\em with} uses a lexical entry that requires it to link to a governing verb to the left, an extracted object farther to the left, and nothing to the right.  Each entry has a hand-assigned cost in \{0,1,2\} and the parser finds the parse of minimum total cost.
% Information about cost is in this message from Linas Vepstas on 2014-02-01: 
%   https://groups.google.com/d/msg/link-grammar/eeJw1Ofgc9U/diqPYSwuFfoJ

@@

\input{figure/link_analysis_table.tex}
